---
title: 'Problem Set #1'
author: "Daniel Kent for w241 Experiments and Causality"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document:
    df_print: paged
  pdf_document: default
---
<!--
Some guidelines for submitting problem sets in this course:

- Please submit a PDF document rather than a Word document or a Google document.
- Please put your name at the top of your problem set.
- If you'll be using `R` or `Python` code to calculate your answers, please put the code and its output directly into your Problem Set PDF document.
- It is highly recommended, although not required, that you use the RMarkdown feature in RStudio to compose your problem set answers. RMarkdown allows you to easily intermingle analysis code and answers in one document. It is of a similar design as `jupyter` and an ipython notebook.
- You do not need to show work for trivial calculations, but showing work is always allowed.
- For answers that involve a narrative response, please feel free to describe the key concept directly and briefly, if you can do so, and do not feel pressure to go on at length.
- Please ask us questions about the problem set if you get stuck. **Don't spend more than 20 minutes puzzling over what a problem means.** 
- Please ensure that someone (us!) can compile your solution set. The best way is to use the web-hosted links we've provided. 
- For text answers, please **embolden** your answers to the questions.
- For computing answers, please store them in the answer objects we have created. 
-->

# 1. Potential Outcomes Notation 

- Explain the notation $Y_{i}(1)$.

**The notation above significes the potential outcome if the ith subject were treated.  As we break it down, Y signifies that it is an outcome, the i subscript refers to the ith subject, and one represents a binary indicator identifying treatment.**

- Explain the notation $E[Y_{i}(1)|d_{i}=0]$.

**The notation above signifies the expected treated potential outcome for the ith subject who does not recieve the treatment when one subject is sampled at random.  Breaking it down, the E[] signifies the expectation of the inside contents when one individual is sampled at random; as stated above, the first term prior to the pipe represents the potential outcome if the ith subject is treated; the pipe signifies the potential outcome when the conents to the right hold, i.e. defining a specific sub-group; and the final $d_{i}=0$ represents that the ith subject does not recieve treatment.  The book describes this as mindbending because these potential outcomes can be imagined but not observed.**

- Explain the difference between the notation $E[Y_{i}(1)]$ and the notation $E[Y_{i}(1)|d_{i}=1]$.
(Extra credit) 

**I believe that the difference relates directly to the pipe, which indicates the criteria that define a subgroup.  So the later term means that the expectation of $Y_{i}(1)$ when one sample is selected at random from those subjects treated, and the former term means th expected value of $Y_{i}(1)$ when one subject is sampled at random, not just from those subjects that were treated.**

- Explain the difference between the notation $E[Y_{i}(1)|d_{i}=1]$ and the notation 
$E[Y_{i}(1)|D_{i}=1]$.  Use exercise 2.7 from FE to give a concrete example of the difference.

**The difference in capital D versus lowercase d is that the capital D represents all possible ways that $d_{i}$ could have been realized, whereas lowercase d represents the outcomes of $d_{i}$ for one experiment.  In exercise 2.7, lowercase d would represent what the exercise suggests - randomly selecting villages 3 and 7 and observing the outcome.  Capital D would represent the outcome if the researcher looked at all possible ways that the villages could have been sampled, that is if they selected villages (1,2)(1,3)(1,4)(1,5)(1,6)(1,7)(2,3)(2,4)(2,5)(2,6)(2,7)(3,4)(3,5)(3,6)(3,7)(4,5)(4,6)(4,7)(5,6)(5,7)(6,7).**

# 2. Potential Outcomes Practice
Use the values in the following table to illustrate that $E[Y_{i}(1)] - E[Y_{i}(0)] = E[Y_{i}(1) - Y_{i}(0)]$.

+---------------+------------+------------+----------+
|               | $Y_{i}(0)$ | $Y_{i}(1)$ | $\tau_i$ |
+===============+============+============+==========+
|Individual 1   |   5        |  6         |  1       |                          
|               |            |            |          |                        
+---------------+------------+------------+----------+
|Individual 2   |   3        |  8         |  5       |                          
|               |            |            |          |                          
+---------------+------------+------------+----------+
|Individual 3   |   10       |  12        |  2       |                          
|               |            |            |          |                          
+---------------+------------+------------+----------+
|Individual 4   |   5        |  5         |  0       |                          
|               |            |            |          |                      
+---------------+------------+------------+----------+
|Individual 5   |  10        |  8         |  -2      |                          
|               |            |            |          |                      
+---------------+------------+------------+----------+

**We are intending to prove that equation of average treatment effects.**


**$\frac{1}{n}\sum\limits^{n}_{i=1}Y_i(1) -\frac{1}{n}\sum\limits^{n}_{i=1}Y_i(0)$**

**$\frac{1}{5}\sum\limits^{5}_{i=1}(6+8+12+5+8) -\frac{1}{5}\sum\limits^{5}_{i=1}(5+3+10+5+10)$**

**$=\frac{1}{5}\sum\limits^{5}_{i=1}(39) -\frac{1}{5}\sum\limits^{5}_{i=1}(33)$**

**$ =\frac{1}{5}\times(39) -\frac{1}{5}\times(33) = 7.8 - 6.6 = 1.2$**

**$\frac{1}{n}\sum\limits^{n}_{i=1}[Y_i(1) - Y_i(0)] $**

**$\frac{1}{5}\sum\limits^{5}_{i=1}(6-5)+(8-3)+(12-10)+(5-5)+(8-10) = \frac{1}{5}\sum\limits^{5}_{i=1}1+5+2+0-2 = \frac{6}{5} = 1.2$**

**$1.2 = 1.2$ Q.E.D**


# 3. Conditional Expectations

Consider the following table:

+---------------+------------+------------+----------+
|               | $Y_{i}(0)$ | $Y_{i}(1)$ | $\tau_i$ |
+===============+============+============+==========+
|Individual 1   |   10       |  15        |  5       |                          
|               |            |            |          |                        
+---------------+------------+------------+----------+
|Individual 2   |   15       |  15        |  0       |                          
|               |            |            |          |                          
+---------------+------------+------------+----------+
|Individual 3   |   20       |  30        |  10      |                          
|               |            |            |          |                          
+---------------+------------+------------+----------+
|Individual 4   |   20       |  15        |  -5      |                          
|               |            |            |          |                      
+---------------+------------+------------+----------+
|Individual 5   |  10        |  20        |  10      |                          
|               |            |            |          |                      
+---------------+------------+------------+----------+
|Individual 6   |  15        |  15        |  0       |                           
|               |            |            |          |                      
+---------------+------------+------------+----------+
|Individual 7   |  15        |  30        |  15      |                          
|               |            |            |          |                      
+---------------+------------+------------+----------+
|Average        |  15        |  20        |  5       |                          
|               |            |            |          |                      
+---------------+------------+------------+----------+

Use the values depicted in the table above to complete the table below. 

\newpage

+---------------+----+----+----+--------------------------+
| $Y_{i}(0)$    | 15 | 20 | 30 | Marginal $Y_{i}(0)$      |
+===============+====+====+====+==========================+
|10             | n:1|n:1 |n:0 |n:2                       |
|               |%:14|%:14|%:0 |%:29                      |
+---------------+----+----+----+--------------------------+
|15             | n:2| n:1| n:1|n:4                       |
|               |%:29|%:14|%:14|%:57                      |
+---------------+----+----+----+--------------------------+
|20             | n:1| n:0| n:1|n:2                       |
|               |%:14|%:0 |%:14|%:29                      |
+---------------+----+----+----+--------------------------+
|Marginal       |n:4 |n:2 |n:2 |n:7                       |
| $Y_{i}(1)$    |%:57|%:29|%:29|%:100                     |
+---------------+----+----+----+--------------------------+

a. Fill in the number of observations in each of the nine cells; 
b. Indicate the percentage of all subjects that fall into each of the nine cells. 
c. At the bottom of the table, indicate the proportion of subjects falling into each category of $Y_{i}(1)$. 
d. At the right of the table, indicate the proportion of subjects falling into each category of $Y_{i}(0)$. 
e. Use the table to calculate the conditional expectation that $E[Y_{i}(0)|Y_{i}(1) > 15]$. 

**In words, I believe it means "find the expectation (i.e. distribution?) that the outcome of the ith subject were not treated $Y_{i}(0)$, given that the ith subject's treatment (i.e.Y_{i}(1)) was greater than 15"**

**So using the marginal distributions, we see that for $Y_{i}(1)$, there are four individual observations that are greater than 15 and now we must take the expectation of these observations:**

$E[X] = \sum x Pr[X=x] = \frac{1}{4}(20)+\frac{1}{4}(20)+\frac{1}{4}(30)+\frac{1}{4}(30)=25$

f. Use the table to calculate the conditional expectation that $E[Y_{i}(1)|Y_{i}(0) > 15]$. 

**In words, I believe it means "find the expectation (i.e. distribution?) that the outcome of the ith subject were  treated $Y_{i}(1)$, given that the ith subject's no treatment (i.e.Y_{i}(1)) was greater than 15"**

**So using the marginal distributions, we see that for $Y_{i}(0)$, there are two individual observations that are greater than 15 and now we must take the expectation of these observations:**

$E[X] = \sum x Pr[X=x] = \frac{1}{2}(15)+\frac{1}{2}(30)=22.5$

# 4. More Practice with Potential Outcomes
Suppose we are interested in the hypothesis that children playing outside leads them to have better eyesight.  

Consider the following population of ten representative children whose visual acuity we can measure.  (Visual acuity is the decimal version of the fraction given as output in standard eye exams.  Someone with 20/20 vision has acuity 1.0, while someone with 20/40 vision has acuity 0.5.  Numbers greater than 1.0 are possible for people with better than "normal" visual acuity.)

```{r, echo = F}
d <- data.frame(child = 1:10, 
                y0 = c(1.1, 0.1, 0.5, 0.9, 1.6, 2.0, 1.2, 0.7, 1.0, 1.1), 
                y1 = c(1.1, 0.6, 0.5, 0.9, 0.7, 2.0, 1.2, 0.7, 1.0, 1.1) )
knitr::kable(d)
```

In the table, state $Y_{i}(1)$ means "playing outside an average of at least 10 hours per week from age 3 to age 6," and state $Y_{i}(0)$ means "playing outside an average of less than 10 hours per week from age 3 to age 6."   $Y_{i}$ represents visual acuity measured at age 6.

a. Compute the individual treatment effect for each of the ten children.  Note that this is only possible because we are working with hypothetical potential outcomes; we could never have this much information with real-world data. (We encourage the use of computing tools on all problems, but please describe your work so that we can determine whether you are using the correct values.)

**For each individual test subject, we know that the way to calculate the causal affect of treatment, $\tau_i$, is the difference between the two potential outcomes:**

$\tau_i = Y_i(1) - Y_i(0)$

**So in this case, we will create a new vector that represents the individual differences for each subject.**
```{r}
d$tau <- with(d, y1-y0)
knitr::kable(d)
```


```{r, echo = F}
answer.POa <- d$tau
```

b. In a single paragraph, tell a story that could explain this distribution of treatment effects.

**A hypothetical story is the following: we identified subjects that participated in the study but to ... varying degrees.  Most subjects played outside for the requisite time and did not see any changes.  Subject 2 took it to the extreme and spent an inordinate amount of time outside, much more than the other subjects, and saw an improvment in eyesight.  With the exception of subject 5, all the other children participated and obeyed in the treatment and did normal "kid" stuff. Subject 2 did go and play outside, but it was reported he brought a smartphone with him and played on it while he was playing outside, in addition to being on it nearly all the other time he was awake or not in class.**

c. What might cause some children to have different treatment effects than others?

**After testing the subjects, we identified two that have experienced a change as a result of treatment.  Subjects 2 and 5 demonstrate individual differences of .5 and -.9, respectively.  We identified progress of subject 2, who has the worst vision of the group without treatment, .1 and possesses vision with treatment, while not perfect, much better than without treatment, .6.  Subject 5's vision has unfortunately decayed without treatment; the vision has declined from without treatment 1.6 to .7 with treatment.  Subject 5 might be explained by general maturity of the eye, as eyes' acuity reduces over time with age, though this is not observed in any of our other subjects and could be representative of an additional lurking variable such as genetics that result in eye accuity deterioration not helped by the treatment (i.e. just bad genes that induced degredation in eye acuity at an advanced young age).  On the other hand, subject 2's vision did improve with treatment to a high degree.  Without more data and context about the individual, we conclude that there could also be lurking variables, leading to the improvment; perhaps the subject has played signficantly more time outside than any of the other subjects, and there is a mysterious threshhold for improvment to occur that the other children didn't achieve.  **

d. For this population, what is the true average treatment effect (ATE) of playing outside.

**Since we have computed the individual treatment effects, to compute the average treatment effect, we need to take a simple mean of the values in the vector.**

```{r}
ate <- mean(d$tau)
ate
```


```{r, echo = F}
answer.POd <- ate
``` 

e. Suppose we are able to do an experiment in which we can control the amount of time that these children play outside for three years.  We happen to randomly assign the odd-numbered children to treatment and the even-numbered children to control.  What is the estimate of the ATE you would reach under this assignment? (Again, please describe your work.)

**To compute this, we would compute the separate ATE **

```{r}

```


```{r, echo = F}
answer.POe <- NA
``` 

f. How different is the estimate from the truth?  Intuitively, why is there a difference?



g. We just considered one way (odd-even) an experiment might split the children. How many different ways (every possible way) are there to split the children into a treatment versus a control group (assuming at least one person is always in the treatment group and at least one person is always in the control group)?

**$\frac{N!}{v!(N-v)!}$, where N = the total number of subjects, v = the number of subjects selected for this draw.  So to do it so there is at least one person always in the treatment group/always in the control group, we vary the v between 1 and 9 and sum up the different equations**

$\frac{10!}{1!(10-1)!}+\frac{10!}{2!(10-2)!}+\frac{10!}{3!(10-3)!}+\frac{10!}{4!(10-4)!}+\frac{10!}{5!(10-5)!}+\frac{10!}{6!(10-6)!}+\frac{10!}{7!(10-7)!}+\frac{10!}{8!(10-8)!}+\frac{10!}{9!(10-9)!}$

```{r}
group <-factorial(10)/(factorial(1)*factorial(10-1))+factorial(10)/(factorial(2)*factorial(10-2))+factorial(10)/(factorial(3)*factorial(10-3))+factorial(10)/(factorial(4)*factorial(10-4))+factorial(10)/(factorial(5)*factorial(10-5))+factorial(10)/(factorial(6)*factorial(10-7))+factorial(10)/(factorial(8)*factorial(10-8))+factorial(10)/(factorial(9)*factorial(10-9))
group
```


```{r, echo = F}
answer.POg <- group
```

**There are 1532 ways to split the children into a treatment versus a control group, assuming at least one person in the treatment group and one in the control group.**

h. Suppose that we decide it is too hard to control the behavior of the children, so we do an observational study instead.  Children 1-5 choose to play an average of more than 10 hours per week from age 3 to age 6, while Children 6-10 play less than 10 hours per week.  Compute the difference in means from the resulting observational data.



```{r, echo = F}
answer.POh <- NA
``` 

i. Compare your answer in (h) to the true ATE.  Intuitively, what causes the difference?


<!--
# 5. FE, exercise 2.5

*Note that the book typically defines D to be 0 for control and 1 for treatment. However, it doesn't have to be 0/1. In particular, one can have more than two treatments, or a continuous treatment variable. Here, the authors want to define D to be the number of minutes the subject is asked to donate. (This is because “D” stands for “dosage.”)*
-->

# 5. Randomization and Experiments

Suppose that a reasearcher wants to investigate whether after-school math programs improve grades. The researcher randomly samples a group of students from an elementary school and then compare the grades between the group of students who are enrolled in an after-school math program to those who do not attend any such program. Is this an experiment or an observational study? Why? 

**We recall that an experiment is an intervention that creates variation in order to teach us about causal questions.  Intervention is the key component of experimentation.  Natural experiments involve naturally occurring data that a researcher argues have the same properties as a true, controlled experiment.  This would require situations where variation was produced by something like random assignment.  Unfortunately the researcher in this example, while randomly samples the group of students int he elementary schools, does not intervene and randomly assign the students to the different treatments.  Consequently this is an observational study since there is no good reason to believe that those who recieved an intervention are otherwise identical to those who did not - perhaps there are other lurking variables like socio-economic-status that facilitate the ability of students to attend after school programs and have other correlated confounds that improve grades like access to tutors, access to nutrious meals and so forth.**

<!--
#7: Skip in 2017
# 7. FE, exercise 2.8
Peisakhin and Pinto report the results of an experiment in India designed to test the effectiveness of a policy called RTIA, which allows citizens to inquire about the status of a pending request from government officials. In their study, the researchers hired confederates, slum dwellers who sought to obtain ration cards (which permit the purchase of food at low cost). Applicants for such cards must fill out a form and have their residence and income verified. The table of results and conditions is below, and there is more in $FE$. 

+--------------------------+-------+------+-----+---------+
|                          | Bribe | RTIA | NGO | Control |
+==========================+=======+======+=====+=========+
| Number Confederates (C)  |    24 |   23 |  18 |      21 |
+--------------------------+-------+------+-----+---------+
| C w/ residence verif     |    24 |   23 |  18 |      20 |
+--------------------------+-------+------+-----+---------+
| M days to verif          |    17 |   37 |  37 |      37 |
+--------------------------+-------+------+-----+---------+
| C w/ ration card 365+    |    24 |   20 |   3 |       5 |
+--------------------------+-------+------+-----+---------+

a. Interpret the apparent effect of the treatments on the proportion of applicants who have their residence verified and the speed with which verification occurred. 
b. Interpret the apparent effect of the treatments on the proportion of applicants who actually received a ration card. 
c. What do these results seem to suggest about the effectiveness of the RTIA as a way of helping slum dwellers obtain ration cards? 
-->


# 6. Lotteries

A researcher wants to know how winning large sums of money in a national lottery affect people's views about the estate tax. The research interviews a random sample of adults and compares the attitudes of those who report winning more than $10,000 in the lottery to those who claim to have won little or nothing. The researcher reasons that the lottery chooses winners at random, and therefore the amount that people report having won is random. 

a. Critically evaluate this assumption. 

****

b. Suppose the researcher were to restrict the sample to people who had played the lottery at least once during the past year. Is it safe to assume that the potential outcomes of those who report winning more than $10,000 are identical, in expectation, to those who report winning little or nothing? 

*Clarifications* 

1. Please think of the outcome variable as an individual's answer to the survey question "Are you in favor of raising the estate tax rate in the United States?"
2. The hint about potential outcomes could be rewritten as follows: Do you think those who won the lottery would have had the same views about the estate tax if they had actually not won it as those who actually did not win it? (That is, is $E[Y_{i}0|D=1] = E[Y_{i}0|D=0]$, comparing what would have happened to the actual winners, the $|D=1$ part, if they had not won, the $Y_{i}(0)$ part, and what actually happened to those who did not win, the $Y_{i}(0)|D=0$ part.) In general, it is just another way of asking, "are those who win the lottery and those who have not won the lottery comparable?"
3. Assume lottery winnings are always observed accurately and there are no concerns about under- or over-reporting.


# 7. Inmates and Reading
A researcher studying 1,000 prison inmates noticed that prisoners who spend at least 3 hours per day reading are less likely to have violent encounters with prison staff. The researcher recommends that all prisoners be required to spend at least three hours reading each day. Let $d_{i}$ be 0 when prisoners read less than three hours each day and 1 when they read more than three hours each day. Let $Y_{i}(0)$ be each prisoner's PO of violent encounters with prison staff when reading less than three hours per day, and let $Y_{i}(1)$ be their PO of violent encounters when reading more than three hours per day. 

In this study, nature has assigned a particular realization of $d_{i}$ to each subject. When assessing this study, why might one be hesitant to assume that ${E[Y_{i}(0)|D_{i}=0] = E[Y_{i}(0)|D_{i}=1]}$ and $E{[Y_{i}(1)|D_{i}=0] = E[Y_{i}(1)|D_{i}=1]}$? In your answer, give some intuitive explanation in English for what the mathematical expressions mean.

**I believe that in this study, the  expressions mean the following:**

${E[Y_{i}(0)|D_{i}=0] = E[Y_{i}(0)|D_{i}=1]}$ **means that the expected probable outcome of violance with the prison staff when reading less than 3 hours per day is equivalent to the expected probable outcome of violance with the prison staff when reading more than 3 hours per day. **


$E{[Y_{i}(1)|D_{i}=0] = E[Y_{i}(1)|D_{i}=1]}$
